---
title: "Pieri Explo survey"
output:
  html_document:
    df_print: paged
  word_document:
    keep_md: yes
---

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
///Installation of packages \\\
--------------------------------------------------------------------------------
--------------------------------------------- -->

```{r setup, include = FALSE, results='hide', message=FALSE, warning=FALSE}
## hide all code chunks in the output, but show errors
knitr::opts_chunk$set(echo = FALSE, error = TRUE, fig.width = 6*1.25, fig.height = 6)


## set default NA to - in output, define figure width/height
options(knitr.kable.NA = "-")



## Installing required packages for this template
required_packages <- c("knitr",       # create output docs
                       "here",        # find your files
                       "rio",         # for importing data
                       "epitrix",     # clean/shape data
                       "dplyr",       # clean/shape data
                       "tidyr",       # clean/shape data
                       "forcats",     # manipulate and rearrange factors
                       "stringr",     # manipulate texts
                       "ggplot2",     # create plots and charts
                       "sitrep",      # MSF field epi functions
                       "survey",      # for survey functions
                       "srvyr"        # dplyr wrapper for survey package
                       )

for (pkg in required_packages) {
  ## install packages if not already present
  if (!pkg %in% rownames(installed.packages())) {
    install.packages(pkg)
  }
  
  ## load packages to this current session 
  library(pkg, character.only = TRUE)
}


## set default text size to 18 for plots
## give classic black/white axes for plots
ggplot2::theme_set(theme_classic(base_size = 18))
```




<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Import Pieri data \\\
--------------------------------------------------------------------------------
--------------------------------------------- -->

```{r import_data, warning= FALSE}
## Excel file ------------------------------------------------------------------

## read in household data sheet
study_data_hh <- rio::import(here::here("1_data", 
                                        "raw", 
                                        "20200815_COMMUNITY_FINDINGS.xlsx"), 
                             which = "General Table",
                             skip = 2, 
                             na = ".") %>% 
  ## convert all variables to lower case
  janitor::clean_names()

## read in individual level data sheet
study_data_indiv <- rio::import(here::here("1_data", 
                                           "raw", 
                                           "20200815_COMMUNITY_FINDINGS.xlsx"), 
                                which = "Detailed HH",
                                skip = 2, 
                                na = ".") %>% 
  ## convert all variables to lower case
  janitor::clean_names()


```



<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// merge_data_levels \\\
--------------------------------------------------------------------------------
For survey datasets with two levels (e.g. household and individual), 
these levels will need to be merged in to one dataset. 

This is done using a unique identifier for the household 
(which has to be repeated for each row in the individuals dataset) 
For a Dharma dataset this variable is "fact_0_id".
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r merge_data_levels, warning= FALSE}

## create and ID variable in both datasets
## HH data
study_data_hh <- study_data_hh %>% 
  mutate(id = paste0(location, "_", hh_number))

## individual data
study_data_indiv <- study_data_indiv %>% 
  mutate(id = paste0(location, "_", hh_number))


## join the individual and household data to form a complete data set
study_data_raw <- study_data_hh %>% 
  left_join(study_data_indiv, by = "id")
  
```



<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// browse_data \\\
--------------------------------------------------------------------------------

You'll want to look at your data. Here are a few ways you can explore.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r browse_data, eval = FALSE}
## view the first ten rows of data
head(study_data_raw, n = 10)

## overview of variable types and contents
str(study_data_raw)

## get summary: 
## mean, median and max values of variables
## counts for categorical variables
## also gives number of NAs
summary(study_data_raw)

## view unique values contained in variables 
## you can run this for any column -- just replace the column name
unique(study_data_raw$sex)
```



<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
This part of the script will create and clean variables in your data.

All your cleaning and variable creation should happen in these chunks.
That way, in case something goes wrong, you can push the small arrow at the top
of the chunk to re-run all the code chunks up to the current one.

The chunks are:
- create_age_group   -- creates the age group variables from age
- factor_vars        -- helps clean factor variables
- numerical vars     -- convert variables to numerical variables

You must adapt this section according to your data!
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// create_age_group \\\ 
--------------------------------------------------------------------------------

This chunk will help you set up your age group variable.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r create_age_group, warning= FALSE}

## Age group variables ----------------------------------------------------------
## This step shows you how to create categorical variables from numeric variables.
## We have some intermediate steps on the way.

## make sure age is an integer 
study_data_cleaned <- study_data_raw %>% 
  mutate(age_months = case_when(
    grepl("M", age) ~ age,
    TRUE ~ NA_character_
  )) %>% 
  mutate(age_months = as.numeric(str_replace(age_months, "M", ""))) %>% 
  mutate(age_years = case_when(
    !grepl("M|N/I", age) ~ as.numeric(age),
    TRUE ~ as.numeric(age_months/12)
  ) )
  

## create an age group variable with 3 age groups 6-11months, 1-4 years, 5-9 years
study_data_cleaned <- study_data_cleaned %>%
  mutate(age_group = factor(case_when(
    age_years > 0 & age_years < 5  ~ "0-4 yrs",
    age_years >= 5 & age_years < 15 ~ "5-14 yrs",
    age_years >= 15 & age_years < 29 ~ "15-29 yrs",
    age_years >= 30 & age_years < 45 ~ "30-44 yrs",
    age_years >= 45 ~ "45+ yrs",
    TRUE ~ NA_character_
  ),
  levels = c("0-4 yrs", 
             "5-14 yrs", 
             "15-29 yrs", 
             "30-44 yrs", 
             "45+ yrs"))
  )

```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// factor_vars \\\
--------------------------------------------------------------------------------

This chunk will help you clean factor variables.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r factor_vars, warning= FALSE}

## tidy up the sex variable and create a new one
study_data_cleaned <- study_data_cleaned %>%
  mutate(sex2 = case_when(
    sex %in% c("F", "W") ~ "Female",
    sex == "M" ~ "Male",
    TRUE ~ NA_character_
  ))

## Violent deaths and natural deaths
study_data_cleaned <- study_data_cleaned %>% 
  ## convert unknown values for violent deaths to the same value
  mutate(num_violent_deaths = as.numeric(case_when(
    number_violent_deaths_in_the_last_3_months_in_the_hh == "Unknown" ~ NA_character_,
    TRUE ~ number_violent_deaths_in_the_last_3_months_in_the_hh
  ))) %>% 
  ## repeat same steps for natural deaths, make all unknowns equal N/I 
  mutate(num_natural_deaths = as.numeric(case_when(
    number_natural_deaths_in_the_last_3_months_in_the_hh == "Unknown" ~ NA_character_,
    TRUE ~ number_natural_deaths_in_the_last_3_months_in_the_hh
  ))) 


## separate out sickness and why didn't access care variables
study_data_cleaned <- study_data_cleaned %>% 
  ## create sickness variable
  mutate(sickness = case_when(
    sickness_why_didnt_access_msf_care %in% c("Abdominal pain",
                                              "AWD + URTI", 
                                              "Chronic condition", 
                                              "Cough", "GSW", "Malaria",
                                              "SRH", "TB", "Other") ~ 
      sickness_why_didnt_access_msf_care,
    TRUE ~ NA_character_
  )) %>% 
  ## create no access care
  mutate(reason_no_care = case_when(
    sickness_why_didnt_access_msf_care == sickness ~ NA_character_,
    sickness_why_didnt_access_msf_care == "N/A" ~ NA_character_,
    sickness_why_didnt_access_msf_care == "Admitted at the clinic" ~ NA_character_,
    TRUE ~ sickness_why_didnt_access_msf_care
  ))


## create a hosting variable - indicates whether a household is hosting
study_data_cleaned <- study_data_cleaned %>% 
  ## if people under and 5 and over being  hosted then give a 1 otherwise 0
  mutate(hosting_household = case_when(
         is.na(number_hosted_members_5y) == F | is.na(number_hosted_members_5y_2) == F ~ 1,
         TRUE ~ 0
         ))

```

<!-- ## create a long version of the dataset to take all the comments into consideration -->
```{r comments_assemble}

comments_overview <- study_data_hh %>% 
  select(main_comment_1:main_comment_5) %>% 
  pivot_longer(cols = main_comment_1:main_comment_5, 
               names_to = "all_comments",
               values_to = "values") %>% 
  mutate(new = values)

```

 <!-- Save comments overview -->

```{r save_comments_doc}
##rio::export(comments_overview, here::here("3_output", "comments_overview.xlsx"))
```




<!-- ### Apply data dictionary cleaning to clean the health facility dataset -->

```{r dict_data_cleaning}

## ## Standardising values --------------------------------------------------------
# Import the dictionary that specifies the values of specific variables
dictionary <- rio::import(here::here("1_data", 
                                     "dictionary_pieri.xlsx"))

## This step fixes the values so that you can read them.
## values like 1/0 will be recoded as "Yes" / "No" based on the dictionary
 comments_overview <- matchmaker::match_df(comments_overview,
  dict  = dictionary,
  from  = "options",
  to    = "values",
  by    = "grp"
 )
```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Convert variables to numeric variables \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r clean_numerical, warning= FALSE}

## Convert number of hosted members variables to be numeric
study_data_cleaned <- study_data_cleaned %>% 
  mutate_at(c("number_hosted_members_5y", "number_hosted_members_5y_2"), as.numeric) %>% 
  ## convert number_time_sick and number_time_visited to numeric variables
  mutate_at(c("number_time_sick_in_the_last_3_months", "number_time_visited_msf_clinic_in_the_last_3_months"), as.numeric)


```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Save data \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->


```{r save_cleaned_data}
# rio::export(study_data_cleaned, str_glue("study_data_cleaned_{Sys.Date()}.xlsx"))
```

# Results

## Survey inclusion 

<!--~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// inclusion_counts \\\
--------------------------------------------------------------------------------

The below chunks calculate values that are displayed with the inline text

Get the counts of clusters and households included 
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~-->

```{r inclusion_counts, warning= FALSE}

## get counts of number of villages
num_clus <- study_data_cleaned %>%
  ## trim data to unique villages
  distinct(location.x) %>% 
  ## get number of rows (count how many unique)
  nrow()

## get counts of number households 
num_hh <- study_data_cleaned %>% 
  ## get unique houses by villages
  distinct(location.x, hh_number.x) %>% 
  ## get number of rounds (count how many unique)
  nrow()

```


`r num_hh` households included across `r num_clus` villages in this analysis. 

<!--~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// cluster_hh_size \\\
--------------------------------------------------------------------------------

The below chunks calculate values that are displayed with the inline text

Get counts of households per cluster and individuals per household. 
From these create medians and standard deviations. 
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~-->

```{r cluster_hh_size, warning= FALSE}

## get counts of the number of households per cluster
clustersize <- study_data_cleaned %>% 
  ## trim data to only unique households within each cluster
  distinct(location.x, hh_number.x) %>%
  ## count the number of households within each cluster
  count(location.x) %>% 
  pull(n)

## get the median number of households per cluster
clustermed <- median(clustersize)

## get the min and max number of households per cluster
## paste these together seperated by a dash 
clusterrange <- str_c(range(clustersize), collapse = "--")

## get counts of children per household 
## do this by cluster as household IDs are only unique within clusters
hhsize <- study_data_cleaned %>% 
  count(location.x, hh_number.x) %>%
  pull(n) 

## get median number of children per household
hhmed <- median(hhsize)
## get the min and max number of children per household
## paste these together seperated by a dash 
hhrange <- str_c(range(hhsize), collapse = "--")

# get standard deviation 
hhsd <- round(sd(hhsize), digits = 1)
```


The median number of households per village was
`r clustermed`, with a range of `r clusterrange`. The median number of individuals
per household was `r hhmed` (range: `r hhrange`, standard deviation: `r hhsd`). 


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Demographic overview of households \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

There was an average of `r study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE) %>% summarise(average_perm5 = round(mean(number_permanent_members_5y, na.rm = TRUE), digits = 0)) %>% pull(average_perm5)` permanent household members under 5 years and an average of `r study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE) %>% summarise(average_perm_over5 = round(mean(number_permanent_members_5y_2, na.rm = TRUE), digits = 0)) %>% pull(average_perm_over5)` permanent household members 5 years and over. There were a total of `r nrow(study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE) %>% filter(hosting_household == 1))` households that were hosting people in their households with an average of `r study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE) %>% summarise(average_host5 = round(mean(number_hosted_members_5y, na.rm = TRUE), digits = 0)) %>% pull(average_host5)` hosted household member under 5 years and `r study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE) %>% summarise(average_host_over5 = round(mean(number_hosted_members_5y_2, na.rm = TRUE), digits = 0)) %>% pull(average_host_over5)` hosted household member over 5 years.


```{r household_overview, warning= FALSE}
## Summary of households
study_data_cleaned %>% 
  distinct(location.x, hh_number.x, .keep_all = TRUE) %>% 
  group_by(location.x) %>% 
  ## calculate average number of violent deaths and range
  summarise(count_household_number = n(),
            mean_permanent_under5 = round(mean(number_permanent_members_5y, 
                                               na.rm = TRUE), digits = 1),
             mean_permanent_over5 = round(mean(number_permanent_members_5y_2, 
                                               na.rm = TRUE), digits = 1),
            mean_host_under5 = round(mean(number_hosted_members_5y, 
                                               na.rm = TRUE), digits = 1),
            mean_host_over5 = round(mean(number_hosted_members_5y_2, 
                                               na.rm = TRUE), digits = 1)) %>% 
  rename("Village"  = location.x,
         "Number households" = count_household_number,
         "Average permanent members <5 yrs" = mean_permanent_under5,
         "Average permanent members >=5 yrs" = mean_permanent_over5,
         "Average hosted members <5 yrs" = mean_host_under5,
         "Average hosted members >= 5 yrs"  = mean_host_over5) %>% 
  gt::gt()
            
            
            
```

## Demographic information


In total we included `r nrow(study_data_cleaned)` individuals in the survey analysis. 

<!--~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// median_age_sex_ratios \\\
--------------------------------------------------------------------------------

The below chunk calculate values that are displayed inline

This returns the median age and sex ratios overall as well as by age group. 
It also pulls the age group with the highest sex ratio. 

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~-->

```{r median_age_sex_ratios, warning= FALSE}
## compute the median age 
medage <- median(study_data_cleaned$age_years, na.rm = TRUE)
## paste the lower and uper quartile together
iqr <- str_c(  # basically copy paste togehter the following
  ## calculate the 25% and 75% of distribution, with missings removed
  quantile(     
    study_data_cleaned$age_years, 
    c(0.25, 0.75), 
    na.rm = TRUE), 
  ## between lower and upper place an en-dash
  collapse = "--")


## compute overall sex ratio 
sex_ratio <- study_data_cleaned %>% 
  count(sex2) %>% 
  pivot_wider(names_from = sex2, values_from = n) %>% 
  mutate(ratio = round(Male/Female, digits = 3)) %>%
  pull(ratio)

## compute sex ratios by age group 
sex_ratio_age <- study_data_cleaned %>% 
  count(age_group, sex2) %>% 
  pivot_wider(names_from = sex2, values_from = n) %>% 
  mutate(ratio = round(Male/Female, digits = 3)) %>%
  select(age_group, ratio)

## sort table by ascending ratio then select the lowest (first)
min_sex_ratio_age <- arrange(sex_ratio_age, ratio) %>% slice(1)

```

Among the `r nrow(study_data_cleaned)` surveyed individuals, there were 
`r fmt_count(study_data_cleaned, sex2 == "Female")` females and 
`r fmt_count(study_data_cleaned, sex2 == "Male")` males . The male to
female ratio was `r sex_ratio` in the surveyed population. The lowest male to
female ratio was `r min_sex_ratio_age$ratio`
in the `r min_sex_ratio_age$age_group` year age group.
The median age of surveyed individuals was `r medage` years (Q1-Q3 of `r iqr`
years). Children under five years of age made up 
`r fmt_count(study_data_cleaned, age_years < 5)` of the surveyed individuals.
The highest number of surveyed indivduals were in the 
`r table(study_data_cleaned$age_group) %>% which.max() %>% names()`
year age group.


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// describe_by_age_group_and_sex \\\
--------------------------------------------------------------------------------

This chunk will create an unweighted table of survey individuals by age group 
and sex.

Note that proportions are of the total sample (not within each gender) 
to change this, set proptotal = FALSE
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r describe_by_age_group_and_sex, warning= FALSE}

tab_linelist(study_data_cleaned, age_group, 
             strata     = sex2, 
             row_total  = TRUE,
             col_total  = TRUE,
             prop_total = TRUE) %>%
  select(- variable) %>%
  rename("Age group" = value) %>% 
  gt::gt()
```


There were `r fmt_count(study_data_cleaned, is.na(sex))` cases missing information on sex and 
`r fmt_count(study_data_cleaned, is.na(age_group))` missing age group.

Age and gender distribution of household population covered by the survey excluding those with missing age values.

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// age_pyramid \\\
--------------------------------------------------------------------------------

This chunk creates an unweighted (using study_data_cleaned) age/sex pyramid
of your cases.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r age_pyramid, warning=FALSE}

plot_age_pyramid(filter(study_data_cleaned, is.na(age_group) == F), 
                 age_group = "age_group", 
                 split_by = "sex2",
                 proportion = TRUE) + 
  labs(y = "Proportion", x = "Age group (years)") + # change axis labels
  theme(legend.position = "bottom",     # move legend to bottom
        legend.title = element_blank(), # remove title
        text = element_text(size = 18)  # change text size
       )
```



## Deaths - violent and natural

There were `r fmt_count(study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE), is.na(num_violent_deaths) == F)` households that had information on the number of violent deaths in the last three months in their household and `r fmt_count(study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE), is.na(num_natural_deaths) == F)` households that provided information on the number of natural deaths in the past three months. In total, there were `r study_data_cleaned %>% distinct(location.x, hh_number.x, .keep_all = TRUE) %>% summarise(total_violent = sum(num_violent_deaths, na.rm = TRUE)) %>% pull(total_violent)` violent deaths reported over the past 3 months.


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Violent deaths in past 3 months \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->
```{r violent_deaths_3months,  warning= FALSE}
## Summary of violent deaths
violent_deaths <- study_data_cleaned %>% 
  distinct(location.x, hh_number.x, .keep_all = TRUE) %>% 
  group_by(location.x) %>% 
  ## calculate average number of violent deaths and range
  summarise(mean_violent_deaths = round(mean(num_violent_deaths, na.rm = TRUE), digits = 1),
            range_violent_deaths = str_c(range(num_violent_deaths, na.rm = TRUE), collapse = "--"))


```

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Deaths due to natural causes in past 3 months \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->
```{r natural_deaths_3months, warning= FALSE}
## Summary of violent deaths
natural_deaths <- study_data_cleaned %>% 
  distinct(location.x, hh_number.x, .keep_all = TRUE) %>% 
  group_by(location.x) %>% 
  ## calculate average number of violent deaths and range
  summarise(mean_natural_deaths = round(mean(num_natural_deaths, na.rm = TRUE), digits = 1),
            range_natural_deaths = str_c(range(num_natural_deaths, na.rm = TRUE), collapse = "--"))


```

There was an average of `r round(mean(study_data_cleaned$num_violent_deaths, na.rm = TRUE), digits = 1)` violent deaths per household in the past 3 months among the households that reported information, with a range of `r str_c(range(study_data_cleaned$num_violent_deaths, na.rm = TRUE), collapse = "--")`. There was an average of `r round(mean(study_data_cleaned$num_natural_deaths, na.rm = TRUE), digits = 1)` natural deaths per household in the past 3 months among the households that reported information, with a range of `r str_c(range(study_data_cleaned$num_natural_deaths, na.rm = TRUE), collapse = "--")`


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Combine tables \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r table_natural_violent_deaths,  warning= FALSE}

violent_deaths %>% 
  left_join(natural_deaths, by = "location.x") %>% 
  rename("Village" = location.x,
         "Average violent deaths" = mean_violent_deaths,
         "Range violent deaths" = range_violent_deaths,
         "Average natural deaths" = mean_natural_deaths,
         "Range natural deaths" = range_natural_deaths) %>% 
  gt::gt()

```


## Sicknesses reported

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Sickness and access to MSF \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r sick_attended_msf, warning= FALSE}

sick_attended_msf <- study_data_cleaned %>% 
  summarise(sum_sickness = sum(number_time_sick_in_the_last_3_months, na.rm = TRUE),
            sum_visited_msf = sum(number_time_visited_msf_clinic_in_the_last_3_months, na.rm = TRUE),
            percent_visited_msf = round((sum_visited_msf/sum_sickness) *100, digits = 0))

```


```{r proportion_triaged_out}


num_people_visited_clinic <- study_data_cleaned %>% 
  ##filter for those who visited_msf_clinic
  filter(number_time_visited_msf_clinic_in_the_last_3_months > 0) %>% 
  tally()

num_people_triaged_out <- study_data_cleaned %>% 
  ##filter for those who visited_msf_clinic
  filter(number_time_visited_msf_clinic_in_the_last_3_months > 0) %>% 
  filter(grepl("Triaged out", reason_no_care)) %>% 
  tally()
  

```

Individuals reported a total of `r sick_attended_msf$sum_sickness` sickness episodes and a total of `r sick_attended_msf$sum_visited_msf` vists to MSF clinics over the past three months. Therefore, individuals reported attending an MSF clinic for `r sick_attended_msf$percent_visited_msf`% of their sickness episodes in the past three months.

Of the `r num_people_visited_clinic %>% pull(n)` people who reporting visiting an MSF clinic one or more times in the past three months, `r num_people_triaged_out %>% pull(n)` reported having been triaged out, which represents `r round((num_people_triaged_out %>% pull(n)/num_people_visited_clinic %>% pull(n)) *100, digits= 0)`%.




<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Sicknesses highlighted \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r proportion_sicknesses, warning= FALSE}
## Calculate number and % of sickness
tab_linelist(study_data_cleaned, sickness) %>% 
  select(-variable) %>% 
  #rename variables
  rename("Sickness" = value) %>% 
  mutate(proportion = round(proportion, digits = 0)) %>% 
  arrange(desc(proportion)) %>% 
  gt::gt()

```

## Reasons for lack of care
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Reason for not accessing health care \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->


```{r no_access_health_care,  warning= FALSE}

 study_data_cleaned %>% 
  ##only those that have been ill in the past 3 months
  filter(number_time_sick_in_the_last_3_months > 0) %>% 
  ## reasons for not accessing care
  tab_linelist(reason_no_care) %>% 
  ## remove variable variable
  select(-variable) %>% 
  ## rename variable
  rename("Reason for lack of care"  = value) %>% 
  ##round proportion
  mutate(proportion = round(proportion, digits = 0)) %>% 
  arrange(desc(proportion)) %>% 
  gt::gt()
  

```


## Comments analysis

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// General comments  \\\
--------------------------------------------------------------------------------

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r general_comments, warning = FALSE}

## Count total number of households
total_households <- nrow(study_data_hh)

## Calculate number of comments and proportion per household
comments_overview %>% 
  filter(new != ".missing") %>% 
  group_by(new) %>% 
  tally() %>% 
  ## add proportion variable
  mutate(proportion = round((n/total_households)*100, digits = 0)) %>% 
  ## tidy up appearance of the new variable
  mutate(new = str_replace_all(new, "_", " ")) %>% 
  mutate(new = str_to_sentence(new)) %>% 
  rename("Comment"  = new) %>% 
  arrange(desc(proportion)) %>% 
  gt::gt()

```



